Evidence for a Third Color Parameter Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2018-07-11 submit/response.txt

It is impossible to verify any claims without making these data public, and
I cannot accept this paper until the data are released publicly.

Response 1:

The data that we use as features are now included in the article.  These are in a new table containing a list of the supernovae, their feature measurements, and uncertainties.

The referee correctly points out that the spectrophotometric time series from which these features are extracted are not yet public. Nevertheless, there is a history in astrophysics for results that could stimulate public interest and activity (e.g. verification of identified trends using independent data sets) to be published before all antecedent data products are made public. We believe that the article fits that category. We believe that the condition of making all the spectroscopy public before publication of this article should be adjudicated by the editor, based on the precedents and standards of ApJ.

@@

The authors do not discuss if their sample is representative of the broader SN Ia sample or the cosmology samples. In fact, the demographics and parameters are not discussed at all, deferring to Chotard et al. Before any referee can properly review this paper (and in addition to the data being released publicly), Chotard et al. must be submitted to a journal and provided to the referee of this paper.

Response 2:

We have added paragraphs in the Data section that describe the selection of our sample.  To address the referees’s specific questions, supernovae of the SN 1991T, SN 1991bg and SN 2002cx subclasses are excluded, one object SN2012c with an SALT2 C parameter much larger than the rest of the sample is excluded, one object whose spectral features make it a $>3\sigma$ outlier in the sample population as described in Leget (2016) is excluded.  The host-galaxy contrast is not found to affect spectral-extraction success when the supernova is at peak brightness.

We refer to previously published articles showing distributions of spectral,
light-curve, and host-galaxy properties of a large subset of our sample.

@@

1. It would be useful to show the covariance matrices for all models and discuss how necessary each measurement is. For instance, the Si EW and x1 are highly correlated, and perhaps only one measurement is required.

Response 3:

We added a table listing the supernovae and the credible intervals of their parameters.

Note that the correlation/covariance between a subset of parameters are already in the article.  For example the correlation matrix of the per-supernova parameters is given and there is a figure of their distribution.  However, there are so many parameters that presenting their covariance matrix within the article is unwieldy.
For example, we did try to insert the 35x35 covariance matrix of the global parameters, but it did not come close to fitting neatly into a manuscript page. We suggest as an alternative that we provide the Monte Carlo chains in the tar archive from which the covariance of any possible parameter pair can be determined.  The chains from the three runs are now available at https://drive.google.com/open?id=1d52nxfF7eS3HblfbmT3fxgiZod_Z-Gjg for the referee’s perusal.  If this is acceptable to the referee and editor, we will move these data products onto SNfactory’s webpage.  We will get a DOI number for this page and include it to the manuscript.

@@

Comment 4:

2. Although there is some discussion about the host mass step, it seems like the authors should have a model with this parameter. That is, the mass step is already used for cosmological analyses and it seems prudent to determine if the new parameter is truly a new color parameter or if it is simply the mass step repackaged.

Response 4:

The purpose of this article is to model supernova colors based on properties directly linked to the supernova.  We then check whether the mass-step could be related to our newly-discovered parameter.  So we do the check but in the opposite (but equivalent) sense that the referee suggests.

### 2018-07-11 submit/fix.tex
After Greg's comments

\added{The current SNfactory sample consists of 198 SNe~Ia
having fully-processed spectrophotometric lightcurve data, including
observations on at least two photometric nights, final references, and a
host spectroscopic redshift. 
These supernovae have well-sampled lightcurves as defined by having
at minimum 5 spectra while still active, with at least four within
epochs  $[-10, 35]$ days, at least one within $[-10, 7]$ days, at least one within $[7, 20]$ days,
and at least one color (Anyone know what this means? See slide 3 of
\url{https://snf-doc.lbl.gov/twiki/pub/Conf/VideoConf20160317/IDR_03-17-2016.pdf}) within $[-8, 10]$ days of peak $B$ brightness.
Synthetic-photometric lightcurves derived from the spectra are fit to the SALT2~(v2.4)
model; we require that supernovae have no more than 20\% points that are
$>0.2$ mag from the best-fit model, and require a Normalized Median Absolute Deviation (nMAD) of residuals
of  $< 0.12$ mag.
%These all have at least 5 spectra while the SN is
%active, and pass the quality cuts suggested by \citet{2010A&A...523A...7G}.
More than 80\% of our SNe are from searches where there was no
pre-selection based on host galaxy properties (those
whose names start with ``SNF'', ``LSQ'', or ``PTF'' in Table~\ref{data:tab}).

@@

Considering that our empirical parameters
are not fundamental supernova parameters, their best-fit values are dependent on sample selection.
Nevertheless, this  article seeks
underlying parameters that significantly
affect SN~Ia colors; any such parameters found in a subsample are still
relevant parameters when dealing with the population as a whole.
Within this context, it is not important whether our sample population is representative
of the underlying population.

@@

Our analysis sample is comprised of the
172
supernovae that have been fully processed and have the data coverage to 
give photometric and spectroscopic statistics described above.
The 
spectroscopic-feature measurements, SALT2 light-curve shape parameters  $x_1$,
and their uncertainties
are presented
in \replaced{\citet{2017Chotard}}{Table~\ref{data:tab}. 
For the subset supernovae for whom it is available, we use the host-galaxy mass given in \citet{2013ApJ...770..108C}.

@@

\colhead{Name} & \colhead{$EW_{\mathrm{Ca}}$ (\AA)} & \colhead{$EW_{\mathrm{Si}}$ (\AA)} & \colhead{$\lambda_{\mathrm{Si}}$ (\AA)} & \colhead{$U$} & \colhead{$B$} & \colhead{$V$} & \colhead{$R$} & \colhead{$I$} & \colhead{$x_1$}

@@

\added{It is important to note that the model and those in subsequent sections are applied to our supernova sample;
the ensuing results thus describe our SN~Ia sample's underlying population. }

@@

\subsection{Results for the Latent Parameter Cofactors  $\vec{\gamma}^0$ and $\vec{\gamma}^1$  }
\label{results1:sec}
\added{The MCMC chain for all parameters are accessible from a DOI link \citep{colordata}.}
In this subsection we concentrate on the results for $\vec{\gamma}^0$ and $\vec{\gamma}^1$.
In our model, $g_0$ and $g_1$ are introduced as latent parameters that are related to linear shifts in band magnitudes through
the  coefficients $\vec{\gamma}^0$ and
$\vec{\gamma}^1$.  Otherwise, no physically-motivated suppositions on the properties of $g$ and $\vec{\gamma}$ are imposed.
Since dust extinction is a dominant determinant of supernova colors that is often described by a 2-parameter linear model, we expect
for our two  color terms to capture the effects of dust.
In this section we reexpress our model $\vec{\gamma}$'s in terms of
the dust-extinction model of \citet{1999PASP..111...63F} \citepalias[henceforth referred to as][]{1999PASP..111...63F}.

@@

\subsection{Results for the Latent Parameter Cofactor $\sigma_p \vec{\phi}$}
\label{results2:sec}
\added{The MCMC chain for all parameters are accessible from a DOI link \citep{colordata}.}
The credible intervals of all Model~II global parameters are given in  \S\ref{m12results:sec}.
In this subsection we focus on the effect of adding the third color term  $\sigma_p \vec{\phi}$.
There is strong evidence for a third color parameter.
The 68\% credible interval for $\sigma_p$ is $0.051^{+0.005}_{-0.004}$~mag, highly incompatible with zero. 
The posterior for $\sigma_p \vec{\phi}$ is shown in Figure~\ref{M1ev:fig}.  
Two maxima are allowed due to the sign degeneracy for $\vec{\phi}$.  Practically each of the MCMC chains converges to one of the solutions and does
not migrate between them by crossing zero.  To aid visibility of this figure we thus impose $\phi_I > 0$.  
 The new parameter does affect colors in a  manner distinct from $g_0$ and $g_1$.
A quadratic fraction of
$0.48^{+0.25}_{-0.17}$ of 
 $\vec{\phi}$ lies perpendicular to the $\vec{\gamma}_0$--$\vec{\gamma}_1$ plane.

@@

\subsection{Results}
\added{The MCMC chain for all parameters are accessible from a DOI link \citep{colordata}.}

\subsubsection{Analysis Method and Validation}

@@

\added{
The correlation matrix in Eq.~\ref{corr:eqn} is not close to being singular, meaning that the each of the features conveys
independent supernova variability.  Nevertheless, a strong correlation between two features means that the determination
of one would reduce the residual dispersion in the other.  The amount of residual dispersion is conveyed in the matrix
\begin{equation}
M =
\begin{pmatrix}
\tiny
\begin{array}{rrrrrrr}
2.76  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.06 \times 10^{ 1} & 2.67  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.05 \times 10^{ 1} & 2.74  \pm 0.06 \times 10^{ 1} & 2.71  \pm 0.06 \times 10^{ 1} & 2.75  \pm 0.06 \times 10^{ 1} \\
7.04  \pm 0.16 \times 10^{ 0} & 7.08  \pm 0.16 \times 10^{ 0} & 7.02  \pm 0.16 \times 10^{ 0} & 3.86  \pm 0.19 \times 10^{ 0} & 7.01  \pm 0.16 \times 10^{ 0} & 6.79  \pm 0.21 \times 10^{ 0} & 7.06  \pm 0.16 \times 10^{ 0} \\
1.98  \pm 0.04 \times 10^{ 1} & 2.03  \pm 0.05 \times 10^{ 1} & 2.05  \pm 0.05 \times 10^{ 1} & 2.03  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} & 2.04  \pm 0.05 \times 10^{ 1} \\
9.92  \pm 0.18 \times 10^{-1} & 5.43  \pm 0.23 \times 10^{-1} & 9.86  \pm 0.19 \times 10^{-1} & 9.95  \pm 0.18 \times 10^{-1} & 9.86  \pm 0.19 \times 10^{-1} & 9.71  \pm 0.26 \times 10^{-1} & 9.92  \pm 0.19 \times 10^{-1} \\
7.33  \pm 0.31 \times 10^{-2} & 7.32  \pm 0.32 \times 10^{-2} & 7.37  \pm 0.32 \times 10^{-2} & 7.32  \pm 0.32 \times 10^{-2} & 7.38  \pm 0.32 \times 10^{-2} & 7.36  \pm 0.33 \times 10^{-2} & 7.36  \pm 0.32 \times 10^{-2} \\
3.82  \pm 0.42 \times 10^{-2} & 3.73  \pm 0.41 \times 10^{-2} & 3.88  \pm 0.43 \times 10^{-2} & 3.79  \pm 0.42 \times 10^{-2} & 3.88  \pm 0.44 \times 10^{-2} & 3.89  \pm 0.44 \times 10^{-2} & 3.88  \pm 0.44 \times 10^{-2} \\
4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.95  \pm 1.09 \times 10^{-2} & 4.97  \pm 1.10 \times 10^{-2} 
\\\end{array}
\end{pmatrix}.
\label{ellipse:eqn}
\end{equation}
For indeces $i, j \in \{EW_{\mathrm{Ca}}, EW_{\mathrm{Si}}, \lambda_{\mathrm{Si}}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), A_{p,V}\}$, $M_{ii}$ is the standard deviation of the parameter $i$ distribution, and for $i \ne j$ $M_{ij}$ is the standard deviation in parameter $i$
if parameter $j$ were known perfectly.  For example, for $i= EW_{\mathrm{Si}}$ the standard deviation is 7.08 but when $j=x_1$
is fixed its
standard deviation drops to 3.86.  Therefore, an analysis that considers $x_1$ but not $EW_{\mathrm{Si}}$ would not account for
$3.86\vec{\beta}=\{0.14 ,  0.09,  0.08,  0.07,  0.04 \}$ ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ magnitude dispersions
that are not intrinsically random.
}

>>>>

\added{The SNfactory sample used here consists of 198 SNe~Ia
having fully-processed spectrophotometric lightcurve data, including
observations on at least two photometric nights, final references, and a
host spectroscopic redshift.   This sample was defined for use in \citet{leget:thesis}.
These supernovae have well-sampled lightcurves as defined by having
at minimum 5 spectra while still active, with at least four within
epochs  $[-10, 35]$ days, at least one within $[-10, 7]$ days, at least one within $[7, 20]$ days,
and at least one color (Anyone know what this means? See slide 3 of
\url{https://snf-doc.lbl.gov/twiki/pub/Conf/VideoConf20160317/IDR_03-17-2016.pdf}) within $[-8, 10]$ days of peak $B$ brightness.
Synthetic-photometric lightcurves derived from the spectra are fit to the SALT2~(v2.4)
model; we require that supernovae have no more than 20\% points that are
$>0.2$ mag from the best-fit SALT2 model, and require a Normalized Median Absolute Deviation (nMAD) of residuals
of  $< 0.12$ mag relative to this model.  \citep[It is known that more flexibility than available to SALT2 is required to capture
the full range of observed SN~Ia behavior, e.g.\ ][]{2011AJ....141...19B, 2011ApJ...731..120M}.
%These all have at least 5 spectra while the SN is
%active, and pass the quality cuts suggested by \citet{2010A&A...523A...7G}.
More than 80\% of our SNe are from searches where there was no
pre-selection based on host galaxy properties (those
whose names start with ``SNF'', ``LSQ'', or ``PTF'' in Table~\ref{data:tab}).

\citet{2013ApJ...770..108C}  present the distribution of host-galaxy parameters for a subset of the  above sample.
\citet{2015MNRAS.447.1247S} show scatter plots of the peak Si~II~$\lambda 5972$ and
Si~II~$\lambda 6355$ velocities and pseudo-equivalent
widths, absolute $B$ and $V$ magnitudes, $B-V$, SALTII $c$ and $x_1$ parameters, and $\Delta m_{15}$ for 
an preceding sample of 151 SNe.

@@

Considering that our empirical parameters
are not fundamental supernova parameters, their best-fit values are dependent on sample selection.
Nevertheless, this  article seeks
underlying parameters that significantly
affect SN~Ia colors; any such parameters found in a subsample are still
relevant parameters when dealing with the population as a whole.
Within this context, it is not so important whether our sample population is representative
of the underlying population.

@@

Our analysis sample is comprised of the
172
supernovae that have been fully processed and have the data coverage to 
give photometric and spectroscopic statistics described above.
The 
spectroscopic-feature measurements, 
\replaced{SALT2 light-curve shape parameters  $x_1$,
and their uncertainties
are presented
in \citet{2017Chotard}}{
synthetic magnitudes (plus a constant offset), SALT2 light-curve shape parameters  $x_1$,
and their uncertainties
are presented
in 
Table~\ref{data:tab}. 
For the subset supernovae for whom it is available, we use the host-galaxy mass given in \citet{2013ApJ...770..108C}.

@@

\colhead{Name} & \colhead{$EW_{\mathrm{Ca}}$ (\AA)} & \colhead{$EW_{\mathrm{Si}}$ (\AA)} & \colhead{$\lambda_{\mathrm{Si}}$ (\AA)} & \colhead{$U$+offset} & \colhead{$B$+offset} & \colhead{$V$+offset} & \colhead{$R$+offset} & \colhead{$I$+offset} & \colhead{$x_1$}

@@

\added{It is important to note that the model and those in subsequent sections are applied to our supernova sample;
the ensuing results thus correspond to our SN~Ia sample's underlying population. }

@@

\subsection{Results for the Latent Parameter Cofactors  $\vec{\gamma}^0$ and $\vec{\gamma}^1$  }
\label{results1:sec}
\added{The MCMC chains for all parameters from this and subsequent sections are accessible from the SNfactory website\footnote{For the referee the parameters
are available at \url{https://drive.google.com/open?id=1d52nxfF7eS3HblfbmT3fxgiZod_Z-Gjg}. Concurrent with publication the chain will be
made available
at \url{https://snfactory.lbl.gov/snf/data/index.html}.}.}
In this subsection we concentrate on the results for $\vec{\gamma}^0$ and $\vec{\gamma}^1$.
In our model, $g_0$ and $g_1$ are introduced as latent parameters that are related to linear shifts in band magnitudes through
the  coefficients $\vec{\gamma}^0$ and
$\vec{\gamma}^1$.  Otherwise, no physically-motivated suppositions on the properties of $g$ and $\vec{\gamma}$ are imposed.
Since dust extinction is a dominant determinant of supernova colors that is often described by a 2-parameter linear model, we expect
for our two  color terms to capture the effects of dust.
In this section we reexpress our model $\vec{\gamma}$'s in terms of
the dust-extinction model of \citet{1999PASP..111...63F} \citepalias[henceforth referred to as][]{1999PASP..111...63F}.

@@

\subsection{Results for the Latent Parameter Cofactor $\sigma_p \vec{\phi}$}
\label{results2:sec}
The credible intervals of all Model~II global parameters are given in  \S\ref{m12results:sec}.
In this subsection we focus on the effect of adding the third color term  $\sigma_p \vec{\phi}$.
There is strong evidence for a third color parameter.
The 68\% credible interval for $\sigma_p$ is $0.051^{+0.005}_{-0.004}$~mag, highly incompatible with zero. 
The posterior for $\sigma_p \vec{\phi}$ is shown in Figure~\ref{M1ev:fig}.  
Two maxima are allowed due to the sign degeneracy for $\vec{\phi}$.  Practically each of the MCMC chains converges to one of the solutions and does
not migrate between them by crossing zero.  To aid visibility of this figure we thus impose $\phi_I > 0$.  
 The new parameter does affect colors in a  manner distinct from $g_0$ and $g_1$.
A quadratic fraction of
$0.48^{+0.25}_{-0.17}$ of 
 $\vec{\phi}$ lies perpendicular to the $\vec{\gamma}_0$--$\vec{\gamma}_1$ plane.

@@

\subsection{Results}
\subsubsection{Analysis Method and Validation}

@@

\added{
The correlation matrix in Eq.~\ref{corr:eqn} is not close to being singular, meaning that the each of the features conveys
independent supernova variability.  Nevertheless, a strong correlation between two features means that the determination
of one would reduce the residual dispersion in the other.  The amount of residual dispersion is conveyed in the matrix
\begin{equation}
Z =
\begin{pmatrix}
\scriptsize
\begin{array}{rrrrrrr}
27.589  \pm 0.554 & 27.451  \pm 0.572 & 26.668  \pm 0.589 & 27.509  \pm 0.542 & 27.408  \pm 0.563 & 27.099  \pm 0.640 & 27.506  \pm 0.562 \\
7.041  \pm 0.160 & 7.076  \pm 0.157 & 7.019  \pm 0.165 & 3.860  \pm 0.186 & 7.011  \pm 0.158 & 6.790  \pm 0.206 & 7.056  \pm 0.159 \\
19.779  \pm 0.449 & 20.295  \pm 0.458 & 20.463  \pm 0.460 & 20.267  \pm 0.456 & 20.430  \pm 0.461 & 20.423  \pm 0.462 & 20.404  \pm 0.465 \\
0.992  \pm 0.018 & 0.543  \pm 0.023 & 0.986  \pm 0.019 & 0.995  \pm 0.018 & 0.986  \pm 0.019 & 0.971  \pm 0.026 & 0.992  \pm 0.019 \\
0.073  \pm 0.003 & 0.073  \pm 0.003 & 0.074  \pm 0.003 & 0.073  \pm 0.003 & 0.074  \pm 0.003 & 0.074  \pm 0.003 & 0.074  \pm 0.003 \\
0.038  \pm 0.004 & 0.037  \pm 0.004 & 0.039  \pm 0.004 & 0.038  \pm 0.004 & 0.039  \pm 0.004 & 0.039  \pm 0.004 & 0.039  \pm 0.004 \\
0.050  \pm 0.011 & 0.050  \pm 0.011 & 0.050  \pm 0.011 & 0.050  \pm 0.011 & 0.050  \pm 0.011 & 0.050  \pm 0.011 & 0.050  \pm 0.011 
\\\end{array}
\end{pmatrix},
\label{ellipse:eqn}
\end{equation}
where the associated units are [\AA, \AA, \AA, 1, mag, mag, mag]$\cdot$[\AA, \AA, \AA, 1, mag, mag, mag]$^{T}$.

For indices $i, j \in \{EW_{\mathrm{Ca}}, EW_{\mathrm{Si}}, \lambda_{\mathrm{Si}}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), A_{p,V}\}$, $M_{ii}$ is the standard deviation of the parameter $i$ distribution, and for $i \ne j$ $Z_{ij}$ is the standard deviation in parameter $i$
if parameter $j$ were known perfectly.  For example, for $i=EW_{\mathrm{Si}}$ the standard deviation is 7.08\AA\ but when $j=x_1$
is fixed its
standard deviation drops to 3.86\AA.  Therefore, an analysis that considers $x_1$ but not $EW_{\mathrm{Si}}$ would not account for
3.86\AA$\vec{\beta}=\{0.14 ,  0.09,  0.08,  0.07,  0.04 \}$~mag ${\hat{U}}{\hat{B}}{\hat{V}}{\hat{R}}{\hat{I}}$ magnitude dispersions
that are not intrinsically random.
}
